---
title: 'Assignment 6: Using Sobol with an ODE'
author: "Katheryn Moya"
output: html_document
date: '2023-05-19'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)

```
# Modeling Dynamic Forest Growth

Consider the following model of forest growth (where forest size in measured in units of carbon (C)):

dC/dt = r ∗ C for forests where C is below a threshold canopy closure
dC/dt = g ∗ (1 − C/K) for forests where carbon is at or above the threshold canopy closure and 
K is a carrying capacity in units of carbon

The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon. Growth rates change from exponential (r) to linear (g) when it reaches the canopy closure threshold.

# Instructions

1. Implement this model in R (as a differential equation)

```{r}
# create and source forest growth model

source("/R/ESM232/esm232_a6/R/dforestgrowth.R") # on remote desktop

# source("~/Documents/MESM Year 2/ESM232/esm232_a6/R/dforestgrowth.R") # on my comp

```


2. Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and using the following parameters:
* canopy closure threshold of 50 kgC
* K = 250 kg C (carrying capacity)
* r = 0.01 (exponential growth rate before before canopy closure)
* g = 2 kg/year (linear growth rate after canopy closure)


```{r}
# set initial forest size 

Cinitial = 10

# come up with first set of parameters
r = 0.01
K = 250
closure = 50
g = 2

#convert to a dataframe
parms <- cbind.data.frame(r, K, closure, g)

# gets results for 300 years (evaluating every year)
years <-seq(from=1, to=300)

# run model for 300 years using ODE solver

results <- ode(y = Cinitial, times = years, func= dforestgrowth, parms = parms)
head(results)
colnames(results)=c("year","C")

# save results as a dataframe for easier plotting
results <- as.data.frame(results)
```

3. Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no uncertainty)


```{r}
ggplot(results, aes(year, C)) +
  geom_line()

```


##4. Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the estimated maximum forest size (e.g maximum of 300 years, varies with these parameters
* pre canopy closure growth rate (r)
* post-canopy closure growth rate (g)
* canopy closure threshold and carrying capacity(K)

Assume that parameters are all normally distributed with means as given above and standard deviation of 10% of mean value

### Create samples to make Sobol Object
```{r}
# reduce output dimensionality by finding the maximum forest size over 300 years

max_c = max(results$C) # maximum of 183.7213 kgC at year 300
## should this be directly in the model?

# we already have our intitial forest size
Cinitial

# obtain parameters sets (from sobel)
# want to learn about sensitivity to pre-canopy closure growth rate (r), post-canopy closure growth rate (g) and carrying capacity (K)

# set the number of parameters
# all parameters are normally distributed w means as given above and sd of 0.1 of mean
np = 200
r = rnorm(mean = 0.01, sd = 0.1, n = np)
g = rnorm(mean = 2, sd = 0.1, n = np)
K = rnorm(mean = 250, sd = 0.1, n = np)
X1 = cbind.data.frame(r = r, g = g, K = K)

# repeat to get second set of samples

r = rnorm(mean = 0.01, sd = 0.1, n = np)
g = rnorm(mean = 2, sd = 0.1, n = np)
K = rnorm(mean = 250, sd = 0.1, n = np)
X2 = cbind.data.frame(r = r, g = g, K = K)

# fix negatives since they aren't meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# create sobol object and get sets of parameters for running model 

sens_P = sobolSalt(model = NULL, X1, X2, nboot = 300)

# parameter sets are
head(sens_P$X)

# add names to columns

colnames(sens_P$X) = c("r", "g", "K")

head(sens_P$X)

```

### Run ODE with sobol object

```{r}
# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

c_wrapper = function (r, g, K, closure, Cinitial, times, func){
  parms = list(r = r, g = g, K = K, closure = closure)
  results = ode(y = Cinitial, times = times, func = func, parms = parms)
  colnames(results) = c("time", "C")
  
  #get metrics
  
  results = as.data.frame(results)
  max_C = max(results$C)
  return(list(max_c = max_c))
}

# use pmap

all_results = as.data.frame(sens_P$X) %>% # order of operations will assume this is the list we are referring to for pmap
  pmap(c_wrapper, closure = closure, Cinitial=Cinitial, times=years, func=dforestgrowth)

# extract out results from pmap into a data frame
allres = all_results %>% map_dfr(`[`,c("max_c"))

```



5. Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two Sobol indices (S and T).

```{r}

# create boxplots
tmp = allres %>% pivot_longer(cols=everything(),names_to="metric", values_to="value")
ggplot(tmp, aes(metric, value, col=metric))+geom_boxplot()

```




6. In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about how what parameters climate change might influence).
Submit R markdown with model implementation, graphs and sensitivity analysis and R file with your model


