---
title: 'Assignment 6: Using Sobol with an ODE'
author: "Katheryn Moya and Sarah Lam"
output:
  html_document:
    code_folding: show
date: '2023-05-19'
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)

```

## Modeling Dynamic Forest Growth

Consider the following model of forest growth (where forest size in measured in units of carbon (C)):

dC/dt = r ∗ C for forests where C is below a threshold canopy closure
dC/dt = g ∗ (1 − C/K) for forests where carbon is at or above the threshold canopy closure and 
K is a carrying capacity in units of carbon

The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon. Growth rates change from exponential (r) to linear (g) when it reaches the canopy closure threshold.


```{r}
# create and source forest growth model

source("/R/ESM232/esm232_a6/R/dforestgrowth.R") # on remote desktop

# source("~/Documents/MESM Year 2/ESM232/esm232_a6/R/dforestgrowth.R") # on my comp

```


## Run model for 300 years with no uncertainty
```{r}
# set initial forest size 

Cinitial = 10

# come up with first set of parameters
r = 0.01
K = 250
closure = 50
g = 2

#convert to a dataframe
parms <- cbind.data.frame(r, K, closure, g)

# gets results for 300 years (evaluating every year)
years <-seq(from=1, to=300)

# run model for 300 years using ODE solver

results <- ode(y = Cinitial, times = years, func= dforestgrowth, parms = parms)
# head(results)
colnames(results)=c("year","C")

# save results as a dataframe for easier plotting
results <- as.data.frame(results)
```


```{r}
# visualize

ggplot(results, aes(year, C)) +
  geom_line(color = "forestgreen") +
  labs(x = "Year",
       y = "Forest Size (kg C)") +
  theme_minimal() 


```


## Run a Sobol global sensitivity analysis for maximum forest size


```{r}
### Create samples to make Sobol Object
# reduce output dimensionality by finding the maximum forest size over 300 years

max_c = max(results$C) # maximum of 183.7213 kgC at year 300


# we already have our intitial forest size
Cinitial

# obtain parameters sets (from sobol)
# want to learn about sensitivity to pre-canopy closure growth rate (r), post-canopy closure growth rate (g) and carrying capacity (K)

# set the number of parameters
# all parameters are normally distributed w means as given above and sd of 0.1 of mean
np = 200
r = rnorm(mean = 0.01, sd = 0.1, n = np)
g = rnorm(mean = 2, sd = 0.1, n = np)
K = rnorm(mean = 250, sd = 0.1, n = np)
X1 = cbind.data.frame(r = r, g = g, K = K)

# repeat to get second set of samples

r = rnorm(mean = 0.01, sd = 0.1, n = np)
g = rnorm(mean = 2, sd = 0.1, n = np)
K = rnorm(mean = 250, sd = 0.1, n = np)
X2 = cbind.data.frame(r = r, g = g, K = K)

# fix negatives since they aren't meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# create sobol object and get sets of parameters for running model 

sens_P = sobolSalt(model = NULL, X1, X2, nboot = 300)

# parameter sets are
head(sens_P$X)

# add names to columns

colnames(sens_P$X) = c("r", "g", "K")

head(sens_P$X)

```


```{r}
### Run ODE with sobol object

# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

c_wrapper = function (results, r, g, K, closure, Cinitial, times, func){
  parms = list(r = r, g = g, K = K, closure = closure)
  results = ode(y = Cinitial, times = times, func = func, parms = parms)
  colnames(results) = c("time", "C")
  
  #get metrics
  
  results = as.data.frame(results)
  max_C = max(results$C)
  return(list(max_C = max_C))
}

# use pmap

all_results = as.data.frame(sens_P$X) # order of operations will assume this is the list we are referring to for pmap|


allres = all_results  %>%  pmap(c_wrapper, results = all_results,closure = closure, Cinitial=Cinitial, times=years, func=dforestgrowth)

# extract out results from pmap into a data frame
final_allres = allres %>% map_dfr(`[`,c("max_C"))

```



## Visualize Sensitivity and Compute Sobol Indices
```{r}
# create boxplots
tmp = final_allres %>% pivot_longer(cols= everything(), names_to="metric", values_to="C")

ggplot(tmp, aes(metric, C, color = metric, fill = metric))+ 
  geom_boxplot(width = 0.3) +
  scale_color_manual(values = "forestgreen") +
  scale_fill_manual(values = "#ccdcc3") +
  theme_minimal() +
  theme(
    legend.position = "none"
  )
  


```


```{r}
## sobol indices - have to do them separately

sens_max_C = sensitivity::tell(sens_P, final_allres$max_C)


#first-order indices (main effect without co-variance)
sens_max_C$S

# total sensitivity index (partitioning the output variance)

sens_max_C$T

```

## Discussion 

The results of our simulation demonstrate that the most variance in forest size is attributed to the exponential growth rate before canopy closure (r). The total effect shows that r and the carrying capacity (K) interact the most with respect to influencing the variance in maximum forest size. This makes sense as the exponential growth rate is limited by the canopy closure threshold, where then forest growth is limited by the carrying capacity. Climate change will further influence these two parameters as decreased water availability may influence growth rates and reduce the overall carrying capacity for forest growth. 

