---
title: 'Assignment 6: Using Sobol with an ODE'
author: "Katheryn Moya"
output: html_document
date: '2023-05-19'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)

```

# Instructions


1. Implement this model in R (as a differential equation)

```{r}
# create and source forest growth model

source("/R/ESM232/esm232_a6/R/dforestgrowth.R")

```


2. Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and using the following parameters:
* canopy closure threshold of 50 kgC
* K = 250 kg C (carrying capacity)
* r = 0.01 (exponential growth rate before before canopy closure)
* g = 2 kg/year (linear growth rate after canopy closure)


```{r}
# set initial forest size 

Cinitial = 10

# come up with first set of parameters
r = 0.01
K = 250
closure = 50
g = 2

#convert to a dataframe
parms <- cbind.data.frame(r, K, closure, g)

# gets results for 300 years (evaluating every year)
years <-seq(from=1, to=300)

# run model for 300 years using ODE solver

results <- ode(y = Cinitial, times = years, func= dforestgrowth, parms = parms)
head(results)
colnames(results)=c("year","C")

results<- as.data.frame(results)
```

# Graph model with no uncertainty

```{r}
ggplot(results, aes(year, C)) +
  geom_line()

```



3. Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no uncertainty)

4. Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the estimated maximum forest size (e.g maximum of 300 years, varies with these parameters
* pre canopy closure growth rate (r)
* post-canopy closure growth rate (g)
* canopy closure threshold and carrying capacity(K)

Assume that parameters are all normally distributed with means as given above and standard deviation of 10% of mean value

5. Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two Sobol indices (S and T).

6. In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about how what parameters climate change might influence).
Submit R markdown with model implementation, graphs and sensitivity analysis and R file with your model




# Dynamic model of Forest Growth 

* models that have feedbacks (conditions evolve through time)

* numerical integration - usually done with a solver 
  
  * only one independent variable ordinary differential equation (e.g just time)
  * we can use the ODE solver
  
  * derivative is first order (e.g $dy/dt$ = ;  not $d^2y/dt$+dy/dy  = f(y,t)) 
  
  * there are ways to solve high order and partial differential equations
